Sparse Function-space Representation \\ of Neural Networks

### 2023-05-14 paper/main.tex
updated CL part

% Math Macros
\newcommand{\MC}{\mbf{C}}
\newcommand{\MZ}{\mbf{Z}}
\newcommand{\MV}{\mbf{V}}
\newcommand{\MX}{\mbf{X}}
\newcommand{\MA}{\mbf{A}}
\newcommand{\MK}{\mbf{K}}
\newcommand{\MI}{\mbf{I}}
\newcommand{\MH}{\mbf{H}}
\newcommand{\T}{\top}
\newcommand{\vzeros}{\mbf{0}}
\newcommand{\vtheta}[0]{\mathbold{\theta}}
\newcommand{\valpha}[0]{\mathbold{\alpha}}
\newcommand{\vkappa}[0]{\mathbold{\kappa}}
\newcommand{\vbeta}[0]{\mathbold{\beta}}
\newcommand{\vlambda}[0]{\mathbold{\lambda}}
\newcommand{\diag}{\text{{diag}}}

@@

%A Gaussian procss characterizes a distribution over functions in terms of the first two moments. The key quantity of 
%A Gaussian process is the prior covariance function (or kernel function) $\kappa(\vx,\vx')$, where $\vx, \vx'$ are input vectors in $\R^{\mathrm{D}}$. We aim to find a kernel based on the neural network to build the Gaussian process posterior. 
%A GP is fully characterized by a mean $\mu(\cdot)$ and covariance function $\kappa(\cdot,\cdot)$ (kernel).
%If we had the following function parameterization $g(\vx) = \phi(\vx)^\top \vw$ where $\phi(\vx)$ is a feature map, we can convert from weight space prior to function space prior giving the following kernel $\kappa(\vx, \vx') = \delta^{-1} \phi(\vx) \phi^\T(\vx')$ (see Ch.~XXX in \cite{rasmussen2006gaussian} for details).\todo{this part needs to be rewritten, starts talking about a general function parameterisation and then jumps directly to the neural network} 
%\subsection{Neural Network to GP Prior} 
%\label{sec:nn2gp}
%
A Gaussian process characterizes a distribution over functions in terms of the first two moments, a mean $\mu(\cdot)$ and covariance function $\kappa(\cdot,\cdot)$ (kernel). We aim to find a kernel and mean function based on the neural network to build the Gaussian process posterior.  A linear function in terms of parameters can be converted from the weight space to the function space (see Ch.~XXX in \cite{rasmussen2006gaussian} as follows:
\begin{equation} \label{eq:weight_func}
g_\weights(\mathbf{x}) = \phi^\top\!(\vx) \, \vw \implies \mu(\vx) = \vzeros \quad \kappa(\mathbf{x}, \mathbf{x}') = \frac{1}{\delta} \phi(\vx)^\top \phi(\vx)
\end{equation}
Given an approximate posterior $q(\vw)$, \citet{immerImprovingPredictionsBayesian2021} proposed to make the following linear approximation
$f^{\textrm{lin.}}_{\vw^*}(\vx; \vw) = f_{\vw^*}(\vx) + \Jac{\weights^*}{\mathbf{x}} \, (\vw - \vw^*)$ of the neural network, where $\Jac{\weights}{\mathbf{x}} \coloneqq \nabla_\weights f(\mathbf{x})^\top \in \R^{C \times P}$ is the Jacobian. The approximation can be used with any approximate inference technique but in the paper they focus on the Laplace-GGN. They find that the linear approximation helps to improve predictions by correcting the approximate posterior when sampled through a non-Gaussian likelihoods.

@@

\section{Sparse function-space sequential learning}

@@

\subsection{Continual learning}

\subsection{Reinforcement learning}
A key challenge in RL is to balance the trade-off between exploration and exploitation.
That is, should the agent select an action that it knows will lead to a high reward (exploitation), or should it
hope to discover new actions which lead to high reward (exploration) by selecting actions that it has not selected before.
In this section, we detail how our method's principled uncertainty estimates can be used to balance this trade-off in a model-based RL algorithim.

We consider stochastic environments with states \(\state \in \stateDomain \subseteq \R^{D_{\state}} \),
actions \(\action \in \actionDomain \subseteq \R^{D_{\action}}\) and transition dynamics
\(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \), such that
$\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}$, where  $\noise_{t}$
is i.i.d. transition noise.
We consider the episodic setting where the system is reset to an initial state $\state_{0}$ at each episode and we
assume that there is a known reward function $r : \stateDomain \times \actionDomain \rightarrow \R$.
% Following the Markov decision process formulation \cite{bellmanMarkovianDecisionProcess1957a}, we
% denote the states \(\state \in \stateDomain \subseteq \R^{D_{\state}} \) and actions \(\action \in \actionDomain \subseteq \R^{D_{\action}}\)
% of the sytem, the reward function $r : \stateDomain \times \actionDomain \rightarrow \R$, and transition dynamics
% \(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \), such that
% $\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}$ where  $\noise_{t}$
% is i.i.d. transition noise.
% $\mathbb{E}_{\noise_{0:\infty}} \big[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \big]$
% \begin{align} \label{eq-model-free-objective}
% \policy^{*} = \arg \max_{\policy \in \policyDomain} J(\transitionFn, \policy) = \arg \max_{\policy \in \policyDomain} \mathbb{E}_{\noise_{0:\infty}} \bigg[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \bigg]
% \quad \text{s.t. } \state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t},
% \end{align}
The goal of RL is to find the policy \(\pi : \stateDomain \rightarrow \actionDomain\)
(from a set of policies $\Pi$) that maximises the sum of discounted rewards
in expectation over the transition noise,
\begin{align} \label{eq-model-free-objective}
J(\transitionFn, \policy) = \mathbb{E}_{\noise_{0:\infty}} \bigg[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \bigg]
\quad \text{s.t. } \state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t},
\end{align}
where $\gamma \in [0, 1)$ is a discount factor.
In this work, we consider model-based RL where a model of the transition dynamics is learned \(f_{\mathbf{w}} \approx \transitionFn\) and then used by a planning algorithm.
A simple approach is to use the learned dynamic model $f_{\mathbf{w}^{*}}$ and maximise the objective in \cref{eq-model-free-objective},
\begin{align} \label{eq-greedy}
  \policy^{\text{Greedy}}(\state) &= \arg \max_{\pi \in \Pi} J(f_{\mathbf{w}^{*}}, \pi).
\end{align}
However, we can leverage the method in \cref{sec:methods} to obtain a function-space posterior over the learned dynamics $q_{\mathbf{u}}(\hat{\transitionFn} \mid \dataset)$,
where $\mathcal{D}$ represents the state transition data set \(\mathcal{D} = \{(s_{i},a_{i}), s_{i+1}\}_{i=0}^{N}\).
Importantly, the uncertainty represented by this posterior distribution can be used to balance the exploration-exploitation trade-off,
using approaches such as posterior sampling \cite{osbandWhyPosteriorSampling2017,osbandMoreEfficientReinforcement2013},
\begin{align} \label{eq-posterior-sampling}
  \policy^{\text{PS}}(\state) &= \arg \max_{\pi \in \Pi} J(\hat{f}, \pi)
\quad \text{s.t. } \tilde{\transitionFn} \sim q_{\mathbf{u}}(\hat{\transitionFn} \mid \dataset),
\end{align}
where a function $\tilde{\transitionFn}$ is sampled from the (approximate) posterior $q_{\mathbf{u}}(\hat{\transitionFn} \mid \dataset)$ and used to find a policy as
in \cref{eq-greedy}.
Intuitively, this strategy will explore where the model has high uncertainty, which in turn will reduce the model's uncertainty as data is collected and used to
train the model.
See \cref{sec-rl-appendix} for more details of how we use

\textbf{Pathwise conditioning in function-space??}
\todo{could try to implement this}

@@

Our experiments seek to answer the following questions:
\begin{itemize}
  \item Does NN2SFR caputure good uncertainty estimates
\end{itemize}

@@

\subsection{Reinforcement Learning}

In section \label{sec-rl} we detailed how our method can be used to

See \label{sec-rl-experiments-appendix} for more details of our experiments.

We compare a model-based RL agent using a simple MLP dynamic model with
\begin{itemize}
  \item uncertainty is good for exploration
\end{itemize}

\cref{fig:rl} shows training curves

@@

\section{Model-based Reinforcement Learning}
\label{sec-rl-appendix}

Model Predictive Path Integral (MPPI) control
\cite{panSample2015}
\cite{williamsModel2017}
\todo{what is correct citation for MPPI?}

% \cref{alg-mbrl} shows the typical model-based RL loop.
\begin{align} \label{eq-fast-update-mpc}
  \policy_{i+1}^{\text{PS}}(\state) = \arg \max_{\action_{0}} \max_{\action_{1:\Horizon}}
\E \bigg[ \sum_{t=0}^{H-1} \gamma^{t} r(\state_{t},\action_{t}) \mid \state_{0}=\state  \bigg] + Q_{\theta}(\state_{\Horizon}, \action_{H})
\quad \text{s.t. } \state_{t+1} &= \hat{\transitionFn}(\state_{t}, \action_{t}) + \noise_{t}
\end{align}
with $\hat{\transitionFn} \sim p(\transitionFn \mid \dataset)$

\subsection{Experiment Configuration}
This section details how we configured and ran our reinforcement learning experiments.

\textbf{Dynamic model}
In all experiments we used an MLP dynamic model with a single hidden layer of width 64 and TanH activation functions.
At each episode we used Adam \cite{adam} to optimise the NN parameters for $5000$ iterations with a learning rate of $0.001$.
We reset the optimizer after each episode.
As we are performing regression we instantiate the loss function in \cref{eq-empirical-risk} as the well-known mean squared error.
This corresponds to a Gaussian likelihood with unit variance.
We then set the prior precision as $\delta=0.0001$.
% It is worth noting that $\delta$ effects both the neur
In all experiments our sparse function-space representation uses $m=128$ inducing points and this seemed to be sufficient.

\textbf{Model predictive path integral (MPPI)}
MPPI is an online planning algorithim which iteratively improves the action trajectory $\action_{t:t+H}$ using samples.
At each iteration $j$, $N=256$ trajectories are sampled according to the currect action trajectory $\action^{j}_{t:t+H}$.
The $K=32$ top trajectories with highest returns $\sum_{h=0}^{H} = r(\state^{j}_{t+h}, \action^{j}_{t+h})$ are selected.
The next action trajectory $\action^{j+1}_{t:t+H}$ is then computed by taking the weighted average of the top $K=32$ trajectories
with weights from the softmax over returns from top $K=32$ trajectories.

$\gamma 0.9$
$\tau=0.005$
horizon $H=5$
temperature $0.5$
momentum $0.1$

\textbf{Initial data set}
We collect an initial data set using a random policy for one episode.

\textbf{DDPG}
action value function is an MLP with a single hidden layer of width $128$ with ELU activation functions.
We train the DDPG agent using Adam for $500$ iterations at each episode, using a learning rate $0.0001$.

>>>>

% Math Macros
\newcommand{\MB}{\mbf{B}}
\newcommand{\MC}{\mbf{C}}
\newcommand{\MZ}{\mbf{Z}}
\newcommand{\MV}{\mbf{V}}
\newcommand{\MX}{\mbf{X}}
\newcommand{\MA}{\mbf{A}}
\newcommand{\MK}{\mbf{K}}
\newcommand{\MI}{\mbf{I}}
\newcommand{\MH}{\mbf{H}}
\newcommand{\T}{\top}
\newcommand{\vzeros}{\mbf{0}}
\newcommand{\vtheta}[0]{\mathbold{\theta}}
\newcommand{\valpha}[0]{\mathbold{\alpha}}
\newcommand{\vkappa}[0]{\mathbold{\kappa}}
\newcommand{\vbeta}[0]{\mathbold{\beta}}
\newcommand{\vlambda}[0]{\mathbold{\lambda}}
\newcommand{\diag}{\text{{diag}}}

@@

%A Gaussian procss characterizes a distribution over functions in terms of the first two moments. The key quantity of 
%A Gaussian process is the prior covariance function (or kernel function) $\kappa(\vx,\vx')$, where $\vx, \vx'$ are input vectors in $\R^{\mathrm{D}}$. We aim to find a kernel based on the neural network to build the Gaussian process posterior. 
%A GP is fully characterized by a mean $\mu(\cdot)$ and covariance function $\kappa(\cdot,\cdot)$ (kernel).
%If we had the following function parameterization $g(\vx) = \phi(\vx)^\top \vw$ where $\phi(\vx)$ is a feature map, we can convert from weight space prior to function space prior giving the following kernel $\kappa(\vx, \vx') = \delta^{-1} \phi(\vx) \phi^\T(\vx')$ (see Ch.~XXX in \cite{rasmussen2006gaussian} for details).\todo{this part needs to be rewritten, starts talking about a general function parameterisation and then jumps directly to the neural network} 
%\subsection{Neural Network to GP Prior} 
%\label{sec:nn2gp}
%

A Gaussian process characterizes a distribution over functions in terms of the first two moments, a mean $\mu(\cdot)$ and covariance function $\kappa(\cdot,\cdot)$ (kernel). We aim to find a kernel and mean function based on the neural network to build the Gaussian process posterior.  A linear function in terms of parameters can be converted from the weight space to the function space (see Ch.~XXX in \cite{rasmussen2006gaussian} as follows:
\begin{equation} \label{eq:weight_func}
g_\weights(\mathbf{x}) = \phi^\top\!(\vx) \, \vw \implies \mu(\vx) = \vzeros \quad \kappa(\mathbf{x}, \mathbf{x}') = \frac{1}{\delta} \phi(\vx)^\top \phi(\vx)
\end{equation}
Given an approximate posterior $q(\vw)$, \citet{immerImprovingPredictionsBayesian2021} proposed to make the following linear approximation
$f^{\textrm{lin.}}_{\vw^*}(\vx; \vw) = f_{\vw^*}(\vx) + \Jac{\weights^*}{\mathbf{x}} \, (\vw - \vw^*)$ of the neural network, where $\Jac{\weights}{\mathbf{x}} \coloneqq \nabla_\weights f(\mathbf{x})^\top \in \R^{C \times P}$ is the Jacobian. The approximation can be used with any approximate inference technique but in the paper they focus on the Laplace-GGN. They find that the linear approximation helps to improve predictions by correcting the approximate posterior when sampled through a non-Gaussian likelihoods.

@@

\section{Sparse Function-space Sequential Learning}

@@

\subsection{Continual Learning}

\begin{equation}
 	\bar{\MB}^{-1}_s = \MKzz^{-1} \vbeta_\vu \MKzz^{-1} \in \R^{m \times m}
 \end{equation}

\begin{equation}
	\mathcal{R_\textit{SFR}}(\mathbf{w}) = \sum_{s=1}^{t-1}\left[\left(f_{\vw}(\MZ_{s}) - f_{\vw_{s}}(\MZ_s) \right)^\T \bar{\MB}^{-1}_{s} \left(f_{\vw}(\MZ_{s}) - f_{\vw_{s}}(\MZ_s) \right) \right] \in \R
\end{equation}

\subsection{Reinforcement learning}
A key challenge in RL is to balance the trade-off between exploration and exploitation.
That is, should the agent select an action that it knows will lead to a high reward (exploitation), or should it select an action that it has
not selected before, which could lead to either a higher or lower reward (exploration).
In this section, we detail how our method's principled uncertainty estimates can be used to balance this trade-off using a model-based RL algorithim.

We consider stochastic environments with states \(\state \in \stateDomain \subseteq \R^{D_{\state}} \),
actions \(\action \in \actionDomain \subseteq \R^{D_{\action}}\) and transition dynamics
\(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \), such that
$\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}$, where  $\noise_{t}$
is i.i.d. transition noise.
We consider the episodic setting where the system is reset to an initial state $\state_{0}$ at each episode and we
assume that there is a known reward function $r : \stateDomain \times \actionDomain \rightarrow \R$.
% Following the Markov decision process formulation \cite{bellmanMarkovianDecisionProcess1957a}, we
% denote the states \(\state \in \stateDomain \subseteq \R^{D_{\state}} \) and actions \(\action \in \actionDomain \subseteq \R^{D_{\action}}\)
% of the sytem, the reward function $r : \stateDomain \times \actionDomain \rightarrow \R$, and transition dynamics
% \(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \), such that
% $\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}$ where  $\noise_{t}$
% is i.i.d. transition noise.
% $\mathbb{E}_{\noise_{0:\infty}} \big[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \big]$
% \begin{align} \label{eq-model-free-objective}
% \policy^{*} = \arg \max_{\policy \in \policyDomain} J(\transitionFn, \policy) = \arg \max_{\policy \in \policyDomain} \mathbb{E}_{\noise_{0:\infty}} \bigg[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \bigg]
% \quad \text{s.t. } \state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t},
% \end{align}
The goal of RL is to find the policy \(\pi : \stateDomain \rightarrow \actionDomain\)
(from a set of policies $\Pi$) that maximises the sum of discounted rewards
in expecation over the transition noise,
\begin{align} \label{eq-model-free-objective}
J(\transitionFn, \policy) = \mathbb{E}_{\noise_{0:\infty}} \bigg[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \bigg]
\quad \text{s.t. } \state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}
\end{align}
where $\gamma \in [0, 1]$ is a discount factor.
In this work, we consider model-based RL where a model of the transition dynamics \(f_{\mathbf{w}} \approx \transitionFn\) is learned and then used by a planning algorithm.
A simple approach is to use the learned dynamic $f_{\mathbf{w}^{*}}$,
\begin{align} \label{eq-fast-update-mpc}
  \policy^{\text{Greedy}}(\state) &= \arg \max_{\pi \in \Pi} J(f_{\mathbf{w}^{*}}, \pi)
\end{align}
However, we can leverage our method to
$p(\transitionFn_{\mathbf{w}^{*}} \mid \dataset) = \mathcal{N}$
\cite{osbandWhyPosteriorSampling2017,osbandMoreEfficientReinforcement2013}
\begin{align} \label{eq-fast-update-mpc}
  \policy^{\text{PS}}(\state) &= \arg \max_{\pi \in \Pi} J(\hat{f}, \pi)
\quad \text{s.t. } \tilde{\transitionFn} \sim q_{\mathbf{u}}(\hat{\transitionFn} \mid \dataset),
\end{align}

Our exact set up uses
Model Predictive Path Integral (MPPI) control
\cite{panSample2015}
\cite{williamsModel2017}
\todo{what is correct citation for MPPI?}
A principled approach
% Given the posterior dynamics \(p(\transitionFn \mid \mathcal{D})\),
% an effective and  common approach is to simply take the expecation over both the aleatoric and epistemic uncertainty,

\textbf{Appendix}

% \cref{alg-mbrl} shows the typical model-based RL loop.
\begin{align} \label{eq-fast-update-mpc}
  \policy_{i+1}^{\text{PS}}(\state) = \arg \max_{\action_{0}} \max_{\action_{1:\Horizon}}
\E \bigg[ \sum_{t=0}^{H-1} \gamma^{t} r(\state_{t},\action_{t}) \mid \state_{0}=\state  \bigg] + Q_{\theta}(\state_{\Horizon}, \action_{H})
\quad \text{s.t. } \state_{t+1} &= \hat{\transitionFn}(\state_{t}, \action_{t}) + \noise_{t} \\
\hat{\transitionFn} &\sim p(\transitionFn \mid \dataset),
\end{align}
with $\hat{\transitionFn} \sim p(\transitionFn \mid \dataset)$

@@

Our experiments seek to answer the following questions:
\begin{itemize}
  \item Does NN2SFR caputure good uncertainty estimates
  \item Can NN2SFR be used to build a summary of the current neural network to mitigate catastrophic forgetting over previous tasks?
\end{itemize}

@@

\subsection{Fast updates in reinforcement learning}

@@

\section{Model-based Reinforcement Learning with Fast Updates}

\subsection{Problem Statement and Background} \label{sec:problem-statement}
We consider environments with states \(\state \in \stateDomain \), actions \(\action \in \actionDomain\) and transition dynamics \(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \) subject to
iid noise \(\noise_{t}\), given by,
\begin{align}
\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}.
\end{align}
\begin{assumption}
  (system properties) The environment's dynamics are $L_{\transitionFn}\text{-Lipschitz}$ continuous and the transition noise $\noise_{t}$ is $\sigma\text{-sub-Gaussian}$ for all $t \geq 0$.
\end{assumption}

\section{Extending the CL regularizer to multi-class settings}

\begin{align}
 	\bar{\MB}^{-1}_s &= \MKzz^{-1} \vbeta_\vu \MKzz^{-1} \in \R^{C \times m \times m}\\
 	& \Kzz \in \R^{C \times m \times m} 
 \end{align}

\begin{equation}
	\mathcal{R_\textit{SFR}}(\mathbf{w}) = \sum_{s=1}^{t-1}\left[\left(f_{\vw}(\MZ_{s}) - f_{\vw_{s}}(\MZ_s) \right)^\T \bar{\MB}^{-1}_{s} \left(f_{\vw}(\MZ_{s}) - f_{\vw_{s}}(\MZ_s) \right) \right] \in \R
\end{equation}
